\chapter{Brownian Motion (Wiener Process)}

\section{Definition}
\textit{The process $W = (W_{t}: t > 0)$ is a \textbf{\color{blue}$\mathbb{P}$-adapted Brownian motion (or $\mathbb{P}$-adapted Wiener process)} if and only if:
\begin{enumerate}
	\item $W_{t}$ is \textbf{\color{blue}continuous} and $W_{0} = 0$
	\item the value of $W_{t}$  is distributed, under $\mathbb{P}$, as a \textbf{\color{blue}normal random variable} $N(0, t)$:
	\begin{eqnarray}
	W_{t} \sim N(0, t)
	\end{eqnarray}
	\item the \textbf{\color{blue}increments} $W_{s + t} - W_{s}$ \textbf{\color{blue}are normally distributed with variance $t - s$} as a $N(0, t)$ under $\mathbb{P}$:
	\begin{eqnarray}
	W_{t} - W_{s} \sim N(0, t - s)
	\end{eqnarray}
	\item \textbf{\color{blue}the increments} $W_{s + t} - W_{s}$ are \textbf{\color{blue}independent} of the behaviour of $W_{r}$ for $r \leq s$:
	\begin{eqnarray}
	W_{s + t} - W_{s} \neq f(W_{r}), \forall r \leq s
	\end{eqnarray}
\end{enumerate}
}\footnote{Through the paper we use $W_{t}$ and $B_{t}$ interchangeably}


\section{Quadratic Variation}
Let $f:[0, T] \rightarrow \mathbb{R}$ be a function the interval $[0, T]$.\\
Let us take partition $\Delta$ of the interval, that:
\begin{eqnarray}
	0 &=& t_{0} < t_{1} < ... < t_{n} = T
\end{eqnarray}
and define its $length$ as:
\begin{eqnarray}
	l(\Delta) &=& \max(t_{i + 1} - t_{i})
\end{eqnarray}
The quadratic variation for the partition $\Delta$ is then
\begin{eqnarray}
	Q(\Delta) &=& \sum_{i = 0}^{n - 1}\Bigg(f\Big(t_{i + 1}\Big) - f\Big(t_{i}\Big)\Bigg)^{2}\geq 0
\end{eqnarray}
The quadratic variation of the function is then the limit of this as the length of the partition goes to zero.

\section{BM properties}


\section{BM Martingales}

\subsection{Martingale}
A stochastic process ${X(t), \forall t \geq 0}$ is a martingale if
\begin{enumerate}
	\item for any $t$ it is integrable (it has a finite expectation):
	\begin{eqnarray}
		\mathbf{E}(X(t)) < \infty
	\end{eqnarray}
	\item and for any $s > 0$
	\begin{eqnarray}
	\mathbb{E}(X(t + s)\vert \mathscr{F}_{t}) &=& X(t)
	\end{eqnarray}
\end{enumerate}

\subsection{BM Martingales}
The list is taken from\\
\textit{"Fima Klebaner. Introduction to Stochastic Calculus with Applications. p.64"}
\begin{itemize}
	\item \textbf{\color{blue}$W_{t}$} - is a martingale
	\item \textbf{\color{blue}$W_{t}^{2} - t$} - is a martingale
	\item for any $u$, \textbf{\color{blue}$e^{uW_{t} - \frac{u^{2}}{2}t}$} - is a martingale
\end{itemize}
\textbf{Proof:} The key idea in establishing the martingale property is that for any function $g$, the conditional expectation of $g(W_{t + s} - W_{t})$ given $\mathscr{F}_{s}$ equals to the unconditional one.

\begin{itemize}
	\item \textbf{\textit{$W_{t}$ is a martingale.}}\\
	We have to show that
	\begin{eqnarray}
	\mathbb{E}(W_{s}\vert\mathscr{F}_{s}) &=& W_{s}, \forall s < t
	\end{eqnarray}
	We can write:
	\begin{eqnarray}
	W_{t} = W_{s} + (W_{t} - W_{s})
	\end{eqnarray}
	Using the linearity of expectation, we have:
	\begin{eqnarray}
	\mathbb{E}(W_{t}\vert \mathscr{F}_{s}) &=& \mathbb{E}(W_{s}\vert \mathscr{F}_{s}) + \mathbb{E}(W_{t} - W_{s}\vert\mathscr{F}_{s})
	\end{eqnarray}
	The first term on the right-hand side is $W_{s}$ since $W_{s}$ contains no information not contained in $\mathscr{F}_{s}$.\\
	Now we show that:\\
	\begin{eqnarray}
	\mathbb{E}(W_{t} - W_{s}) &=& 0
	\end{eqnarray}
	Recall that
	\begin{eqnarray}
	(W_{t} - W_{s}) \sim N(0, t - s)
	\end{eqnarray}
	As $W_{t} - W_{s}$ is independent of the value of $W_{s}$ and the path up to time $s$, so when we condition on information available at time $s$, we are effectively conditioning on no information. This means that:
	\begin{eqnarray}
	\mathbb{E}[(W_{t} - W_{s})\vert \mathscr{F}_{s}]&=& \mathbb{E}(W_{t} - W_{s}) = 0
	\end{eqnarray}
	Thus BM is a martingale.
	\item Process in the form:
	\begin{eqnarray}
	dX_{t} &=& \sigma(X, t)dW_{t}
	\end{eqnarray}
	is a martingale.
\end{itemize}

\section{TODO}
\begin{itemize}
	\item BM construction
	\item Joshi1.page155 - creation of martingales 
\end{itemize}


\section{Problems}
\begin{itemize}
	\item is $Z = \sqrt{t}N(0,1)$ a BM?
	\item is $X_{t} = \rho W_{1} + \sqrt{1 - \rho^{2}}\tilde{W_{t}}$ a BM?
\end{itemize}
